Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-09-26 RNARedPrint.tex

We introduced a novel approach, based on a general framework and efficient algorithms, to design RNA
sequences, while targeting very specific complex properties; in particular, we describe the targeting of the free energies of multiple target strucutures and the \GCb-content.
Our method combines an FPT stochastic sampling algorithm
with multi-dimensional Boltzmann sampling over distributions
controlled by expressive RNA energy models. Compared to the
previously best available sampling method (uniform sampling), this
approach generated significantly better seed sequences for instances
of an extensive multi-target design benchmark, including pseudoknots.
%
To substantiate our work additionally, we establish the $\#${\sf P}-hardness of uniform sampling;
from a complexity-theoretic point of view, this motivates the FPT, tree decomposition-based nature of our method.

The presented method enables new possibilities for the sequence generation
in the field of RNA sequence design, allowing to enforce additional
constraints, like \GCb\%, while controlling the energy of
multiple target structures. Thus it presents a major advance over
previously applied ad-hoc sampling and even efficient uniform sampling
strategies. We demonstrated the practical feasibility of such controlled sequence
generation, including the case of pseudoknotted structures, and studied its use for multi-target RNA design. Moreover,
our framework is extensible to include various more complex sequence
constraints, including mandatory/forbidden motifs at specific
positions, or anywhere in the designed sequences, by adapting formal language constructs
introduced by \citet{Zhou2013}. In future work, negative design principles could 
be explicitly supported at the generation stage, for instance by
penalizing a set of alternative helices/structures.
{\em Add RNASKetch to conclusion for more complex design scenario + need for generic constratins-based framework for flexibility}

>>>>

We introduced a novel approach, based on a general framework and efficient algorithms, to design RNA
sequences, while targeting very specific complex properties. In particular, we describe the targeting of the free energies of multiple target strucutures and the \GCb-content.
Our method combines an FPT stochastic sampling algorithm
with multi-dimensional Boltzmann sampling over distributions
controlled by expressive RNA energy models. Compared to the
previously best available sampling method (uniform sampling), this
approach generated significantly better seed sequences for instances
of an extensive multi-target design benchmark, including pseudoknots.
%
To substantiate our work additionally, we establish the $\#${\sf P}-hardness of uniform sampling which,
from a complexity-theoretic point of view, motivates the FPT, tree decomposition-based nature of our method.

The presented method enables new possibilities for the sequence generation
in the field of RNA sequence design, allowing to enforce additional
constraints, like \GCb\%, while controlling the energy of
multiple target structures. Thus it presents a major advance over
previously applied ad-hoc sampling and even efficient uniform sampling
strategies. We demonstrated the practical feasibility of such controlled sequence
generation, including the case of pseudoknotted structures, and studied its use for multi-target RNA design.

%perspectives:
Our framework is extensible to include various more complex sequence constraints, including mandatory/forbidden motifs at specific
positions, or anywhere in the designed sequences, by adapting formal language constructs
of Zhou \emph{et al}~\citet{Zhou2013}. In future work, negative design principles could 
be explicitly supported at the generation stage, for instance by
penalizing a set of alternative helices/structures.

{Stefan: \em Mention RNAsketch to perspectives for more complex design scenario + need for generic constraints-based framework for flexibility ($\to$ generic constraint framework).}

{Yann: \em Connection with statistics (design as perfect random sequences for null-model), mention Rivas/Eddy lncRNAs papers...}
